\section{Introduction}\label{sec:introduction} % (fold)
In this draft of the material i am going to have to dispense with the
usual writing conventions adopted in papers on these topics. i'm going
to have adopt whatever tone i need at the time i'm writing up the
calculations. Sometimes this may be very conversational; others it may
be the barest mathematical grunts; others still it may be that i have
lifted text from one of my other papers because the exposition of some
point was better said there. i hope that my readers are not unduly put
out by this decision. i'm not doing this to flout convention or be
rebellious. i find these calculations very technically challenging. To
keep everything going technically, something has to give; i have to
let go of some cognitive burden. So, the academic writing style --
with all of its trade-offs in terms of facilitating technical
communication -- is what i'm letting go of. Perhaps subsequent drafts
can be tightened and polished, but for now, i'm going to speak as if
we were sitting together in a coffee shop with a laptop, wifi and a
pad of paper and a pencil.

So, here's what i have to say. We -- you and i, comfortably ensconced
in our coffee shop and well-equipped with our tools -- can realize and
carry out the calculations of quantum mechanics over a very different
formal theory of dynamics, a formal theory of dynamics that
corresponds to a theory of concurrent computation with
\emph{reflection}. It has the advantage that the underlying theory is
already `quantized', but supports analogues all of the continuuous
operations. Strikingly, this underlying theory has recently been
connected with a notion of metric that we can show, by calculating
together, coincides with the metric induced by the inner product.

There are a lot of reasons why you might be interested in seeing
calculations of this form. Here's why i'm interested. For the past
several centuries there has been no competitor to the ``Newtonian''
account of dynamics. As a result the predominant share of accounts of
dynamical systems and situations have had to be formulated in terms of
the Newtonian machinery. i view this as an intellectually dangerous
position to occupy. Everything, despite it's intrinsic shape, turns
into a nail to be hit with this hammer. Recently, however, the theory
of computation has matured to the point where we have candidates for
theories of dynamics that offer very different perspective on
reasoning about dynamical systems and situations. Testing these
candidates against very successful accounts of dynamical situations,
like quantum mechanics, is going to give us some sense of how mature
they are and some measure of the quality of these accounts of
dynamics.

\subsection{Summary of contributions and outline of paper}

So, we're going to develop an interpretation of the operations of
quantum mechanics normally interpreted by Hilbert spaces and
operators. We're going to do this over a theory of computation. Note
that this is very different than the usual quantum computation program
which develops notions of computation over quantum mechanics. Rather,
we are developing a story that aligns with Wheeler's slogan: It from
Bit. To do this we will first provide an account of the theory of
computation at play here. Then we will dive into a calculation-driven
interpretation of the operations of quantum mechanics.

The reason we take this approach is that -- until very recently --
there hasn't been an axiomatic account of quantum mechanics. As a
result there has been no sharp delineation of the mathematical theory
supporting interpretation of the physical theory and the physical
theory, itself. So, ambient features of the maths are free to be
exploited (or supressed) without a real accounting of their physical
relevance. There is no sharp statement ``here's the physical theory''
qua \emph{theory} and ``here's the mathematical interpretation''
enabling a judgment of how faithful the interpretation is -- apart
from experimental observation. When there is an axiomatic account we
can judge how well a given mathematical formalism supports an
interpretation of the axioms, independent of
experimentation. Likewise, we can judge how well we have captured our
physical evidence and experience with our axiomatics, independent of
any specific mathematical implementation, with accidental detail that
may or may not have physical significance. 

In lieu of a fully fleshed out and vetted axiomatic account of quantum
mechanics, interpreting the operational notions in service of modeling
physical systems will have to suffice. In other words, we are not in
the business of providing a model of Hilbert spaces and operators. We
are in the business of providing a model of quantum mechanics because
we are motivated by testing our notions of dynamics against physical
theory; and, the predictive calculations of the physical theory must
serve as the best formulation -- shy of a fully fleshed out axiomatic
account -- of the physical theory itself (as they have for scientific
theories since time immemorial). Put another way, despite a
whole-hearted commitment to an It-from-Bit ontology, we are firmly
aligned with the shut-up-and-calculate camp as the best way to obtain
results either from the physical perspective or as a quality assurance
measure of our fledgling theory of dynamics.

In detail, we present a reflective process calculus. Then we develop
intuitive correspondences between the notions available in this
calculus and the usual physical notions supporting quantum mechanical
calculations. Thus, 

\begin{table}[htp]
  \center{
    \fbox{
      \begin{tabular}{c|c}
        quantum mechanics & process calculus \\
        \hline
        scalar & name \\
        state vector & process \\
        dual & contextual duals \\
        matrix & formal sums of process-context-dual pairs \\
        orthogonality & process annihilation \\
        inner product & execution-formula + quoting
      \end{tabular}
    }
  }
  \caption{QM - process calculi correspondences}
\end{table}

Then we tighten up these intuitions to operational definitions. We
employ the Dirac notation as the best proxy we can find for an
abstract syntax of the quantum mechanical notions. The definitions we
develop put us in contact with equational constraints coming from the
theory that we demonstrate the definitions and calculations satisfy.

This puts us in a position to shut up and calculate for the
Stern-Gerlach experimental set up, showing how these predictive
calculations become calculations on processes in our theory of a
reflective process calculus.

Penultimately, we demonstrate that the notion of metric coming from
the inner product coincides with the notion of metric available from
the theory of bisimulation. This demonstration gives us the right to
think of space as arising from behavior. Finally, we consider where we
might go from the new vantage point we have obtained.

% section introduction (end)